# Models and effect sizes

Summarizing a relationship with a function

We'll start with models with a single *degree of flexibility*, that is $^\circ{\cal F} = 1$. This includes all the settings covered in most introductory stats courses.

Degrees of flexibility more generally. Explain why "degrees of freedom" is harder than degress of flexibility. You can count degrees of flexibility.

## Activity 2: Showing relationships

Each of the following settings shows n = 100 data points. The variables are shown on the axes and, sometimes, as the name of a color legend. For each of the settings, your task is to:

1. Describe the relationship depicted. Be as quantitative as one could expect  for an estimate *by eye*.
2. Give the name of the appropriate inferential technique.
    - Do you teach that technique in your introductory course.
    - If not, where might students encounter that technique.
3. Attempt to estimate, *by eye* and without whatever supplemental arithmetic is requires, the approximate values of the quantities produced by the technique in (2). For instance, is the relationship between the response and explanatory variables statistically discernable.^[Again, I'm using a proposed replacement for the traditional "significant."] 

### Setting A

```{r echo  = FALSE}
set.seed(101)
CPS <- CPS85 %>% filter(wage < 30) %>% sample_n(size  = 100) %>%
    mutate(sexF = as.numeric(sex == "F"),
           marriedY = as.numeric(married == "Married"))
```

```{r echo = FALSE}
resA <- make_model_plots(marriedY ~ sex, data = CPS, 
                         show_model = TRUE)
resA$P1 %>%
    gf_refine(
        scale_y_continuous(
            'Marital status', breaks  =  c(0, .25, .5, .75, 1), 
            labels = c("Single", .25,  .5, .75, "Married"),
            limits = c(-.1, 1.1)))
```

### Setting B

```{r echo = FALSE}
resB  <- make_model_plots(wage ~ union, data = CPS,
                       show_model = TRUE,
                       sd  =  TRUE, R2 = TRUE,
                       F = TRUE)
resB$P1
```

### Setting C

```{r echo = FALSE}
resC  <- make_model_plots(marriedY ~ age, data = CPS,
                       show_model = TRUE,
                       sd  =  TRUE, R2 = TRUE,
                       F = TRUE)
resC$P1  %>%
    gf_refine(
        scale_y_continuous(
            'Marital status', 
            breaks  =  c(0, .25, .5, .75, 1), 
            labels = c("Single", .25, .5, .75, "Married"), 
            limits = c(-.1, 1.1)))
```

### Setting D

```{r echo = FALSE}
resD  <- make_model_plots(wage ~ educ, data = CPS,
                       show_model = TRUE,
                       sd  =  TRUE, R2 = TRUE,
                       F = TRUE)
resD$P1 
```

### Setting E

```{r echo = FALSE}
resE  <- make_model_plots(marriedY ~ age + sex, data = CPS,
                       show_model = TRUE,
                       sd  =  TRUE, R2 = TRUE,
                       F = TRUE)
resE$P1  %>%
    gf_refine(
        scale_y_continuous(
            'Marital status', 
            breaks  =  c(0, .25, .5, .75, 1), 
            labels = c("Single", .25, .5, .75, "Married"), 
            limits = c(-.1, 1.1)))
```

### Setting F

```{r echo = FALSE}
resF  <- make_model_plots(wage ~ educ* sex, data = CPS,
                       show_model = TRUE,
                       sd  =  TRUE, R2 = TRUE,
                       F = TRUE)
resF$P1 
```

### Setting G

```{r echo = FALSE}
resG <- make_model_plots(marriedY ~ sex  *  union, data = CPS, 
                         show_model = TRUE)
resG$P1 %>%
    gf_refine(
        scale_y_continuous(
            'Marital status', breaks  =  c(0, .25, .5, .75, 1), 
            labels = c("Single", .25,  .5, .75, "Married"),
            limits = c(-.1, 1.1)))
```


### Setting H

```{r echo = FALSE}
resH <- make_model_plots(marriedY ~ sector, data = CPS, 
                         show_model = TRUE)
resH$P1 %>%
    gf_refine(
        scale_y_continuous(
            'Marital status', breaks  =  c(0, .25, .5, .75, 1), 
            labels = c("Single", .25,  .5, .75, "Married"),
            limits = c(-.1, 1.1)))
```

### Setting EYE

```{r echo = FALSE}
resI <- make_model_plots(wage ~ sector, data = CPS, 
                         show_model = TRUE)
resI$P1 
```

### Setting J

```{r echo = FALSE}
resJ <- make_model_plots(wage ~ sector  *  educ, data = CPS, 
                         show_model = TRUE,)
resJ$P1 
```

