---
title: "Activity: Graphical settings for inference"
subtitle: "JMM Minicourse: Stats for Data Science"
author: "Daniel Kaplan"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
SDSdata::sds_setup()
library(LittleApp)
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, 
                      cache.extra = packageVersion('tufte'),
                      out.width="80%", fig.align="center", 
                      fig.width  = 6, fig.height  = 3,
                      eval=TRUE)
options(htmltools.dir.version = FALSE)
questions<-" inferential technique:\\newline \\newline inferential quantities:"
show_model  <- TRUE
show_results <- function(res) {
  #res$P1
  #anova(res$model)
  with(res, 
       gridExtra::grid.arrange(P1, P2$P, P2$label, 
                               nrow = 1, 
                               widths = c(3,1,1)))
}
```

Each of the following settings shows n = 100 data points. The variables are shown on the axes and, sometimes, as the name of a color legend. For each of the settings, your task is to:

1. Describe the relationship depicted. Be as quantitative as one could expect  for an estimate *by eye*.
2. Give the name of the appropriate inferential technique.
    - Do you teach that technique in your introductory course.
    - If not, where might students encounter that technique.
3. Attempt to estimate, *by eye* and without whatever supplemental arithmetic is requires, the approximate values of the quantities produced by the technique in (2). For instance, is the relationship between the response and explanatory variables statistically discernable.^[Again, I'm using a proposed replacement for the traditional "significant."] 
4. Save this sheet. When we get to  the section on model functions, I'm going to ask  you to  draw an  appropriate model function for each setting.

```{r child="settings.Rmd"}
```

