


```{r echo  = FALSE,  eval=TRUE}
set.seed(101)
CPS <- CPS85 %>% filter(wage < 30) %>% sample_n(size  = 100) %>%
    mutate(sexF = as.numeric(sex == "F"),
           marriedY = as.numeric(married == "Married"))
```

`r tufte::newthought("Setting A")``r margin_note(questions)`
```{r echo = FALSE}
resA <- make_model_plots(married ~ sex, data = CPS,
                         show_model = show_model)
show_results(resA)
```

`r tufte::newthought("Setting B")``r margin_note(questions)`
```{r echo = FALSE}
resB  <- make_model_plots(wage ~ union, data = CPS,
                       show_model = show_model)
show_results(resB)
```

\newpage

`r tufte::newthought("Setting C")``r margin_note(questions)`
```{r echo = FALSE}
resC  <- make_model_plots(married ~ age, data = CPS,
                       show_model  =  show_model)
show_results(resC)
```

`r tufte::newthought("Setting D")``r margin_note(questions)`
```{r echo = FALSE}
resD  <- make_model_plots(wage ~ educ, data = CPS,
                       show_model  =  show_model)
show_results(resD)
```

`r tufte::newthought("Setting E")``r margin_note(questions)`
```{r echo = FALSE}
resE  <- make_model_plots(married ~ ns(age,2) * sex,
                          data = CPS,
                          show_model  =  show_model)
show_results(resE)
```

`r tufte::newthought("Setting F")``r margin_note(questions)`
```{r echo = FALSE}
resF  <- make_model_plots(wage ~ educ* sex, data = CPS,
                       show_model  =  show_model)
show_results(resF)
```

`r tufte::newthought("Setting G")``r margin_note(questions)`
```{r echo = FALSE}
resG <- make_model_plots(married ~ sex  *  union, data = CPS,
                         show_model  =  show_model)
show_results(resG)
```


`r tufte::newthought("Setting H")``r margin_note(questions)`
```{r echo = FALSE}
resH <- make_model_plots(married ~ sector, data = CPS,
                         show_model  =  show_model)
show_results(resH)
```

`r tufte::newthought("Setting I")``r margin_note(questions)`
```{r echo = FALSE}
resI <- make_model_plots(wage ~ sector, data = CPS,
                         show_model  =  show_model,
                         R2 = FALSE,  F = FALSE)
show_results(resI)
```

`r tufte::newthought("Setting J")``r margin_note(questions)`
```{r echo = FALSE}
resJ <- make_model_plots(wage ~ sector  *  educ, data = CPS,
                         show_model  =  show_model,
                         R2 = FALSE,  F  =  FALSE)
show_results(resJ)
```

